---
title: "Lab 02:  Ice duration and air temperature in Madison, WI"
subtitle: "Inference for simple linear regression"
author: "Dav King"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

## Setup

Load packages and data:

```{r load-packages}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
library(knitr)
```

```{r load-data}
#| message: false
#| warning: false

ice <- read_csv("data/wi-icecover.csv")
air <- read_csv("data/wi-air-temperature.csv")
```

\pagebreak

## Exercises

### Exercise 1

```{r ex-1}
icecover_avg <- ice %>% 
  group_by(lakeid, year) %>% 
  summarize(avg_ice_duration = mean(ice_duration))
```

There are 270 observations of 3 variables in `icecover_avg`.

\pagebreak

### Exercise 2

```{r ex-2}
airtemp_avg <- air %>% 
  group_by(year) %>% 
  summarize(avg_air_temp = mean(ave_air_temp_adjusted))

ice_air_joined <- inner_join(icecover_avg, airtemp_avg)
```


\pagebreak

### Exercise 3

```{r ex-3}
ggplot(ice_air_joined, aes(x = avg_ice_duration)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Average Number of Days", y = "Number of Observations",
       title = "Distribution of Days between Freeze and Breakup") +
  theme(plot.title = element_text(hjust = 0.5))

fivenum(ice_air_joined$avg_ice_duration)
mean(ice_air_joined$avg_ice_duration)
sd(ice_air_joined$avg_ice_duration)
```

These data are roughly unimodal and normally distributed, with a center around 100 (mean = 99.049, sd = 17.426, median = 101) and a spread from 21 to 140, with first and third quartiles at 91 and 111, respectively. There appears to be a possible low-end outlier at 21 - otherwise, the data all seem to be clustered closely with one another and there are not serious concerns about outliers.

\pagebreak

### Exercise 4

```{r ex-4}
ggplot(ice_air_joined, aes(x = avg_air_temp)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(title = "Distribution of Average Annual Air Temperature",
       x = "Temperature in Degrees Celsius", y = "Number of Years") +
  theme(plot.title = element_text(hjust = 0.5))

fivenum(ice_air_joined$avg_air_temp)
mean(ice_air_joined$avg_air_temp)
sd(ice_air_joined$avg_air_temp)
```

These data are roughly unimodal and normally distributed, with a center around 7.5 (mean = 7.595, sd = 0.98, median = 7.486) and a spread from 4.811 to 10.760, with first and third quartiles at 6.978 and 8.155, respectively. There appear to be a few outliers below an annual average of 5 degrees or above an average of 10; otherwise, there are no notable outliers in the data.

\pagebreak

### Exercise 5

```{r ex-5}
ggplot(ice_air_joined, aes(x = year, y = avg_ice_duration, color = lakeid)) +
  geom_line(alpha = 0.7) +
  labs(x = "Year", y = "Average Number of Days", color = "Lake",
       title = "Number of Days Between Freeze and Breakup by Year",
       subtitle = "Based on Two Wisconsin Lakes") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

Over time, the average number of days seems to have been trending slightly lower in both of the two lakes. However, there is also a lot of variability from year to year, and the trend can only be seen over a long period of time.

\pagebreak

### Exercise 6

```{r ex-6}
ggplot(ice_air_joined, aes(x = year, y = avg_air_temp)) +
  geom_line() +
  labs(x = "Year", y = "Average Temperature in Degrees Celsius",
       title = "Average Air Temperature by Year in Madison, Wisconsin") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

Over time, the average air temperature seems to be increasing over time. However, it still fluctuates from year to year - the trend can only be seen when looking at a long period of time.

\pagebreak

### Exercise 7

$$avg\_ice\_duration = \beta_0 + \beta_1 \times avg\_air\_temp$$

\pagebreak

### Exercise 8

```{r ex-8}
airModel <- linear_reg() %>% 
  set_engine("lm") %>% 
  fit(avg_ice_duration ~ avg_air_temp, data = ice_air_joined)

airModel %>% 
  tidy() %>% 
  kable(digits = 3)
```

$$\hat{avg\_ice\_duration} = 149.454 - 6.636 \times avg\_air\_temp $$

The slope of -6.636 means that, for every one degree increase in annual temperature, we would expect to see a decrease of 6.636 days between freeze and breakup for the ice on the lakes. The intercept does have a theoretically meaningful interpretation, since it is possible (though highly unlikely) to have a year with an average temperature of 0; in such a year, we would expect an average of 149.454 days between freeze and breakup.

\pagebreak

### Exercise 9

$H_0: \beta_1 = 0$, or the slope of the relationship between average air temperature and the number of days between freeze and breakup in a given year is equal to zero.

$H_a: \beta_1 \neq 0$, or the slope described above is not equal to zero.

```{r ex-9}
set.seed(9)

perm_fits <- ice_air_joined %>% 
  specify(avg_ice_duration ~ avg_air_temp) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  fit()

perm_fits %>% 
  filter(term == "avg_air_temp") %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Slope Estimate", y = "Number of Observations",
       title = "Distribution of Slope Estimates in Null Distribution") +
  theme(plot.title = element_text(hjust = 0.5))

obs <- ice_air_joined %>% 
  specify(avg_ice_duration ~ avg_air_temp) %>% 
  fit()

get_p_value(
  perm_fits,
  obs_stat = obs,
  direction = "two-sided"
)
```

The p-value of getting a statistic like ours if there is no relationship between the two variables is ~0. Thus, we can reject $H_0$. We have sufficient evidence to suggest that there is a significant relationship between average air temperature and the number of days between freeze and breakup in any given year.

\pagebreak

### Exercise 10

```{r ex-10}
set.seed(10)

nullCI <- ice_air_joined %>% 
  specify(avg_ice_duration ~ avg_air_temp) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  fit()

nullCI %>% 
  filter(term == "avg_air_temp") %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram(color = "white") +
  theme_bw() +
  labs(x = "Slope Estimate", y = "Number of Observations",
   title = "Distribution of Bootstrapped Samples of the Air Temperature Slope")+
  theme(plot.title = element_text(hjust = 0.5))

get_confidence_interval(
  nullCI,
  point_estimate = obs,
  level = .95,
  type = "percentile"
)
```

The confidence interval says that we can be 95% confident that the true population slope of the relationship between average air temperature and average ice duration lies on the interval (-8.62, -4.64). This is consistent with the results of the hypothesis test because the confidence interval does not contain zero - if it did, we would not be able to reject the null hypothesis that there is no relationship between the two variables.